---
title: "Cyclist_proj1022"
output: html_document
date: "2022-10-12"

---

```{r}
#Data Load In : 
cyclist <- read.csv("C:/Users/Kareena/Desktop/Personal Project/Casestudy_cyclisticOct22/Data/Divvy_Trips_2020_Q1.csv")

#Taking a look at data 
#head(cyclist) - 13 variables, 426887 observations 

```

## Exploratory Data Analysis 
```{r}
# Summarizing Data
summary(cyclist)

#Dimensions of Data 
dim(cyclist)


#Visualizations:
library(ggplot2)
startedAt <- cyclist$start_station_id
start <- hist(startedAt,main="Frequency of Cyclist Start Stations",
xlab="Start Station Id's")
text(start$mids,start$counts,labels = start$counts,adj = c(0.5,-0.5))

EndedStat <- cyclist$end_station_id
end <- hist(EndedStat,main="Frequency of Cyclist End Stations",
xlab="End Station Id's")
text(end$mids,end$counts,labels = end$counts,adj = c(0.5,-0.5))

```


The summary() function showed the min, 1st quartile,median,mean,3rd quartile, Max and NA's of each category variable. 
The variables in the data set are: Ride_id, rideable_type,started_at,ended_at,start_station_name,start_station_id,end_station_name, end_station_id, start_lat,start_lng,end_lat,end_lng,and member_casual 

# Histograms
Initially interested in looking at the difference between the start station and end station id. I wanted to understand if cyclist started and finished at different cyclist stations. 
## In the first histogram: Frequency of Cyclist Start Station Id's 
We observe that there is a higher frequency of stations that cyclist start their route with the ids (0-100) in comparison to those station Id's that are in the range of (350-700). 
##The second histogram: Frequency of Cyclist End Station Id's 
We observe a similar trend of a higher frequency of end stations are used with the ids (0-100) than those with ids in the range of (350-700).

#Why Is this?



```{r pressure, echo=FALSE}
#Data Transformation:
cyclist$start_date <- as.Date(cyclist$started_at)
#Convert the time from 24-hour to a 12-hour time 
cyclist$start_time <- format(as.POSIXct(cyclist$started_at) ,format = "%H:%M:%S")
cyclist$start_time <- format(as.POSIXct(cyclist$start_time ,format = "%H:%M:%S"),format ="%I:%M:%S %p")

cyclist$end_date <- as.Date(cyclist$ended_at)

#Convert the time from 24-hour to a 12-hour time 
cyclist$end_time <- format(as.POSIXct(cyclist$ended_at),format = "%H:%M:%S")

cyclist$end_time <- format(as.POSIXct(cyclist$end_time,format = "%H:%M:%S"),format ="%I:%M:%S %p")


library(tidyverse)
library(lubridate)
#separate(cyclist,start_date,c("Start_Year","Start_Month","Start_Day"),sep = " ")
#separate(cyclist,end_date,c("End_Year","End_Month","End_Day"),sep = "-")

#Separates the start date year, month, day 
cyclist2 <- separate(cyclist,start_date,  c("start_year","start_month","start_day"))
#Separates the End date year, month, day 
cyclist3 <- separate(cyclist,end_date,  c("end_year","end_month","end_day"))



# Plotting the start date 
sapply(cyclist2,class)
StartedMonth <- as.numeric(cyclist2$start_month)
startM <- hist(StartedMonth,main="Frequency of Cyclist Start Months",
xlab="Months ")
text(startM$mids,startM$counts,labels = startM$counts,adj = c(0.6,-0.06))

#Plotting the End Date 
sapply(cyclist3,class)
EndedMonth <- as.numeric(cyclist3$end_month)
EndM <- hist(EndedMonth,main="Frequency of Cyclist End Months",
xlab="Months ")
text(EndM$mids,EndM$counts,labels = EndM$counts,adj = c(0.6,-0.6))


```

#Data Transformation Explained 
Start at and End at
 - Current Format date(YYYY/MM/DD) and Time 
 - New format date needs its own column and time needs to be 

