---
title: "Cyclist_proj1022"
output: html_document
date: "2022-10-12"

---

```{r}
#Data Load In : 
cyclist <- read.csv("C:/Users/Kareena/Desktop/Personal Project/Casestudy_cyclisticOct22/Data/Divvy_Trips_2020_Q1.csv")

#Taking a look at data 
#head(cyclist) - 13 variables, 426887 observations 

```

## Exploratory Data Analysis 
```{r}
# Summarizing Data
summary(cyclist)

#Dimensions of Data 
dim(cyclist)


#Visualizations:
library(ggplot2)
jpeg("freqstartidHist.jpg")
startedAt <- cyclist$start_station_id
start <- hist(startedAt,main="Frequency of Cyclist Start Stations",
xlab="Start Station Id's")
text(start$mids,start$counts,labels = start$counts,adj = c(0.5,-0.5))
dev.off()

jpeg("freqendIdHist.jpg")
EndedStat <- cyclist$end_station_id
end <- hist(EndedStat,main="Frequency of Cyclist End Stations",
xlab="End Station Id's")
text(end$mids,end$counts,labels = end$counts,adj = c(0.5,-0.5))
dev.off()

```


The summary() function showed the min, 1st quartile,median,mean,3rd quartile, Max and NA's of each category variable. 
The variables in the data set are: Ride_id, rideable_type,started_at,ended_at,start_station_name,start_station_id,end_station_name, end_station_id, start_lat,start_lng,end_lat,end_lng,and member_casual 

# Histograms
Initially interested in looking at the difference between the start station and end station id. I wanted to understand if cyclist started and finished at different cyclist stations. 
## In the first histogram: Frequency of Cyclist Start Station Id's 
We observe that there is a higher frequency of stations that cyclist start their route with the ids (0-100) in comparison to those station Id's that are in the range of (350-700). 
##The second histogram: Frequency of Cyclist End Station Id's 
We observe a similar trend of a higher frequency of end stations are used with the ids (0-100) than those with ids in the range of (350-700).





```{r pressure, echo=FALSE}
#Data Transformation:
cyclist$start_date <- as.Date(cyclist$started_at)
#Convert the time from 24-hour to a 12-hour time 
cyclist$start_time <- format(as.POSIXct(cyclist$started_at) ,format = "%H:%M:%S")
cyclist$start_time <- format(as.POSIXct(cyclist$start_time ,format = "%H:%M:%S"),format ="%I:%M:%S %p")

cyclist$end_date <- as.Date(cyclist$ended_at)

#Convert the time from 24-hour to a 12-hour time 
cyclist$end_time <- format(as.POSIXct(cyclist$ended_at),format = "%H:%M:%S")

cyclist$end_time <- format(as.POSIXct(cyclist$end_time,format = "%H:%M:%S"),format ="%I:%M:%S %p")


library(tidyverse)
library(lubridate)
#separate(cyclist,start_date,c("Start_Year","Start_Month","Start_Day"),sep = " ")
#separate(cyclist,end_date,c("End_Year","End_Month","End_Day"),sep = "-")

#Separates the start date year, month, day 
cyclist2 <- separate(cyclist,start_date,  c("start_year","start_month","start_day"))
#Separates the End date year, month, day 
cyclist3 <- separate(cyclist,end_date,  c("end_year","end_month","end_day"))



# Plotting the start date (Month)
sapply(cyclist2,class)
jpeg("StartMonthFreq.jpeg")
StartedMonth <- as.numeric(cyclist2$start_month)
startM <- hist(StartedMonth,main="Frequency of Cyclist Start Months",
xlab="Months ",labels = TRUE)
dev.off()


#Plotting the End Date (Month)
jpeg("EndMonthFreq.jpeg")
sapply(cyclist3,class)
EndedMonth <- as.numeric(cyclist3$end_month)
EndM <- hist(EndedMonth,main="Frequency of Cyclist End Months",
xlab="Months ", labels = TRUE)
dev.off()



```

#Data Transformation Explained 
Start at and End at
 - Current Format date(YYYY/MM/DD) and Time 
 - New format date needs its own column and time needs to be 


#Investigating 
- How are annual member and casual member trips different? 
i. (are their trips shorter or longer?)
ii. (Which stations do annual members use more frequently)?
iii. (Which stations do casual users use more frequently)?

```{r}
# Sorting out casual members: 
library(dplyr)

#an_members <- cyclist[order(cyclist$member_casual),]
#head(an_members)
annual_mem <- filter(cyclist, member_casual =="member")

# Cacluting distance of trips from longitute and latitude of all annual  members 
# This section will include the use of the package geosphere 
library(geosphere)
annual_mem %>% rowwise%>% mutate(distance=distm(x=c(start_lng,start_lat),y=c(end_lng,end_lat)))

#Sorting out annual members 
casual_mem <- filter(cyclist,member_casual =="casual")

# Cacluting distance of trips from longitute and latitude of all casual  members 
casual_mem %>% rowwise %>% mutate(distance = distm(x = c(start_lng,start_lat),y=c(end_lng,end_lat)))




```

